# Networked Contours

## Load General Libraries
```{r}
library(tidyverse)
library(sf)
library(ggmap)
library(tidygraph)
library(igraph)
```

## Define Global Variables
```{r}
# define the path to the dss files
dss_path <- "~/Documents/nrel/data/sfo/solar_extreme_batteries_high_timeseries"
csv_path <- "~/Documents/nrel/data/sfo/solar_extreme_batteries_high_timeseries_csv"
var_path <- "10x/variablesdss"
regions <- list.dirs(csv_path, full.names = FALSE, recursive = FALSE)
region <- "P13U"
```

## Read in Data
```{r}
# read in RDS
lines.df <- readRDS(paste(csv_path, region, "lines_df.rds", sep = "/"))
buses.df <- readRDS(paste(csv_path, region, "buses_df.rds", sep = "/"))
tx.df <- readRDS(paste(csv_path, region, "tx_df.rds", sep = "/"))
```

## Conduct Network Analysis
- Calculate the distance matrix for the network --- yikes this might be too big
- Recalculate the voltages at each node

## Clean the Data to be Ready to Build the Network
```{r}
# rename bus1 as "from" and bus2 to "to" in lines.df and tx.df
lines.df <- lines.df %>% rename(from = bus1, to = bus2, name = line)
tx.df <- tx.df %>% rename(from = bus1, to = bus2)
nodes.df <- buses.df %>% rename(name = bus)

glimpse(nodes.df)
glimpse(lines.df)
glimpse(tx.df)

# combine
edges.df <- bind_rows(lines.df, tx.df)

# check for nodes in neither edge list
nodes.df %>% filter(!(name %in% edges.df$from | name %in% edges.df$to)) %>% nrow()

# filter out any nodes that are not in the edges "from" or "to"
nodes.df <- nodes.df %>%
  filter(name %in% edges.df$from | name %in% edges.df$to)
# from 296286 to 296286 > none removed!!
# export
saveRDS(edges.df, paste(csv_path, region, "edges_df.rds", sep = "/"))
saveRDS(nodes.df, paste(csv_path, region, "nodes_df.rds", sep = "/"))
```

## Build the Network
```{r}
nodes.df <- readRDS(paste(csv_path, region, "nodes_df.rds", sep = "/"))
edges.df <- readRDS(paste(csv_path, region, "edges_df.rds", sep = "/"))


# ensure all edges have a node
graph <- tidygraph::tbl_graph(
  nodes = nodes.df,
  edges = tibble(edges.df),
  directed = FALSE
)
graph
saveRDS(graph, paste(csv_path, region, "graph.rds", sep = "/"))
```

## Check the Number of Components in the Network
```{r}
# read in the network
graph <- readRDS(paste(csv_path, region, "graph.rds", sep = "/"))

# calculate components
graph <- graph %>%
  activate(nodes) %>%
  mutate(component = components(graph)$membership)

# add a column that counts the number of nodes in each component
graph <- graph %>%
  group_by(component) %>%
  mutate(n = n()) %>%
  ungroup()

# show components with the most number of nodes
graph %>% activate(nodes) %>% as_tibble() %>% count(component) %>% arrange(desc(n)) %>% nrow()
graph %>% activate(nodes) %>% as_tibble() %>% count(component) %>% arrange(desc(n)) %>% head(10)
# Round 1: 1 component with 80,644 nodes, then everything else is less than 150 nodes per component
# Round 2: 1 component with 80,915 nodes, then everything else is less than 150 nodes per component

# inspect things
# graph
# max(graph %>% activate(nodes) %>% as_tibble() %>% pull(component)) # 46k components, can't work with that

# # count unique values of number of nodes in each component
graph %>% activate(nodes) %>% as_tibble() %>% count(component) %>% count(n)%>% print(n=50)

# # show me a histogram of the number of nodes in each component
# # use log scale
# graph %>% activate(nodes) %>% as_tibble() %>% count(component) %>% ggplot(aes(x = n)) + geom_histogram(binwidth = 1) + scale_y_log10()
 
saveRDS(graph, paste(csv_path, region, "graph.rds", sep = "/"))
```

      n buses/    n components w/
      component   this many buses in it
      <int>      <int>
 1        2       9188
 2        3       1527
 3        4       2528
 4        5       3115
 5        6       2845
 6        7       2337
 7        8       2268
 8        9       2110
 9       10       1938
10       11       1498
11       12       1330
12       13        924
13       14        771
14       15        484
15       16        332
16       17        206
17       18        131
18       19         75
19       20         37
20       21         16
21       22          7
22       23          2
23       24          2
24       25          1
25       27          1
26       28          1
27       51          1
28       80          1
29      148          1
30    80915          1

## Join with Buses.SF and Plot
```{r}
nodes.df <- readRDS(paste(csv_path, region, "nodes_df.rds", sep = "/"))
graph <- readRDS(paste(csv_path, region, "graph.rds", sep = "/"))

# add the "component" and "n" columns to nodes.df
nodes.df <- nodes.df %>%
  left_join(graph %>% activate(nodes) %>% as_tibble() %>% select(name, component, n), by = c("name" = "name"))

glimpse(nodes.df)
write.csv(nodes.df, paste(csv_path, region, "nodes_w_comp_df.csv", sep = "/"))

# keep just buses where n > 1
nodes.df.filt <- nodes.df %>%
  filter(n > 1000)


# plot the buses
component_plot <- nodes.df.filt %>%
  ggplot() +
  geom_sf(aes(color = as.factor(component)), size = 0.5) +
  theme_bw() 
  # theme(legend.position = "none")

ggsave(
  filename = "10x/figures/component_plot_2.png", 
  component_plot,
  width = 12960,
  height = 12960,
  dpi = 300,
  units = "px",
  device = "png",
  limitsize = FALSE
)
```











## Force Directed Network Layout FR
```{r}
# library(igraph) # for network analysis
# # read in graph
# graph <- readRDS(paste(csv_path, region, "graph.rds", sep = "/"))

# # create force directed layout
# fr <- as.data.frame(layout_with_fr(graph, dim = 2))
# ggplot(fr, aes(x = V1, y = V2)) +
#   geom_point(alpha = 0.2, size = 0.5) +
#   theme_bw()
```

## Run Network Analysis
```{r}
library(igraph) # for network analysis
# read in graph and inspect
graph <- readRDS(paste(csv_path, region, "graph.rds", sep = "/"))
graph
##############################################
# calculate the distances between all nodes
dist_matrix <- distances(graph, mode = "all")
##############################################
# calculate a similarity matrix as 1 / (h+1)^2
sim2_matrix <- 1 / ((dist_matrix + 1)^2)

##############################################
# export matrices
saveRDS(dist_matrix, "10x/variablesdss/dist_matrix.rds")
saveRDS(sim2_matrix, "10x/variablesdss/sim2_matrix.rds")

# remove variables from RAM until needed
rm(dist_matrix, sim2_matrix)
```

## Proceed with Weighted Network Analysis
```{r}
# import buses.sf
buses.sf <- readRDS("10x/variablesdss/buses_sf.rds")
sim2_matrix <- readRDS("10x/variablesdss/sim2_matrix.rds")

# set up a function to calculate weighted voltage considering n nearest neighbors by hops (including self)
neighbors <- 3

# function to weight voltage
weight_voltage <- function(voltage, row) {
  weights <- sim2_matrix[, row]
  indices <- order(weights, decreasing = TRUE)[1:neighbors]
  vw <- data.frame(
    voltages = buses.sf$voltage[indices],
    weights = weights[indices]
  )
  vw$weighted_voltage <- vw$voltages * vw$weights
  sum_weights <- sum(vw$weights)
  vw$scaled_weighted_voltage <- vw$weighted_voltage / sum_weights
  weighted_voltage <- sum(vw$scaled_weighted_voltage)
  return(weighted_voltage)
}
##############################################
# Apply weighted voltage function to buses.sf
Sys.time()
buses.sf <- buses.sf %>% mutate(
  weighted_voltage_n3_d2 = purrr::map_dbl(
    row_number(),
    ~ weight_voltage(voltage, .x)
  )
)
Sys.time()

# filter out any columns that contain "_dev_"
# buses.sf <- buses.sf %>% select(-contains("_dev_"))

# sort the columns by name
# buses.sf <- buses.sf %>% select(sort(names(.)))

# reexport buses.sf
saveRDS(buses.sf, "10x/variablesdss/buses_sf.rds")
```



# Create Contours

## Load Contour Map Libraries
```{r}
library(terra)
library(gstat)
```

## Create Contours
```{r}
# load in lines buffer
buses.sf <- readRDS("10x/variablesdss/buses_sf.rds")
lines.buffer <- readRDS("10x/variablesdss/lines_buffer_100m.rds")

# Convert the lines buffer to a vector
lines.buffer.sv <- vect(lines.buffer)
# get the extent of lines.buffer.sv
ext <- ext(lines.buffer.sv)
# create a grid of the same extent
grd <- rast(ext, res=0.0001)

# convert buses.sf to buses.df
buses.df <- buses.sf %>% st_drop_geometry()
# drop nas from any column
buses.df <- na.omit(buses.df)
# rename lng and lat to x and y
names(buses.df) <- c("name", "voltage", "x", "y")
# glimpse(buses.df)

neighbors <- 1000
# define interpolation model
mg <- gstat(
  id = "idw", 
  formula = voltage~1, 
  locations = ~x+y, 
  data=buses.df, 
  nmax=neighbors, 
  set=list(idp = 2)
)
######################
contour <- interpolate(grd, mg, debug.level=0, index=1)
saveRDS(contour, "10x/variablesdss/contour_n1000_rast.rds")
######################
# plot(contour)
contour.mask <- mask(contour, lines.buffer.sv)
saveRDS(contour.mask, "10x/variablesdss/contour_mask_n1000_rast.rds")
# plot(contour.mask)

# export the contour to RDS
# class(contour) # [1] "SpatRaster", attr(,"package"), [1] "terra"

# clear out extra variables not needed later
rm(buses.sf, buses.df, lines.buffer, lines.buffer.sv, ext, grd, neighbors, mg)
```

## Prepare Contour for Plotting
```{r}
# read in contour object
contour.mask <- readRDS("10x/variablesdss/contour_mask_n1000_rast.rds")

# convert to data.frame
contour.df <- as.data.frame(contour, xy=TRUE)

# rename idw.pred to voltage
contour.df <- contour.df %>% rename(voltage = idw.pred)

# remove rows with NA in any column
# contour.df <- na.omit(contour.df) # removes none

# export df 
saveRDS(contour.df, "10x/variablesdss/contour_n1000_df.rds")
glimpse(contour.df)

rm(contour)
```
Rows: 3,553,912
Columns: 3
$ x        <dbl> -122.3447, -122.3446, -122.3445, -122.3444, -122.3443, -122.3…
$ y        <dbl> 37.90743, 37.90743, 37.90743, 37.90743, 37.90743, 37.90743, 3…
$ voltage <dbl> 0.9800596, 0.9800594, 0.9800592, 0.9800590, 0.9800588, 0.9800…

## Plot Contours
```{r}
# import map
map <- readRDS("10x/variablesdss/map_stadia_z14.rds")
contour.df <- readRDS("10x/variablesdss/contour_n1000_df.rds")

# glimpse(contour.df)
# plot it
contour_plot <- ggmap(map) +
  geom_raster(
    data = contour.df,
    aes(x=x, y=y, fill = voltage),
    alpha = 1.0,
    inherit.aes = FALSE
  ) +
#   geom_sf(
#     data = lines.sf, 
#     color="black", 
#     alpha = 0.05, 
#     inherit.aes = FALSE
#   ) +
  scale_fill_gradient2(
    low = "darkblue",
    mid = "white",
    high = "darkred",
    midpoint = 1,
    limits = c(0.95, 1.05),
    space = "Lab",
    guide = "colourbar",
    aesthetics = "fill",
    oob = scales::squish
  ) +
  theme(
    legend.position = c(0.96, 0.15),
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.x = element_blank(),
    axis.ticks.y = element_blank(),
    text = element_text(family = "Serif", size = 18)
  ) +
  coord_sf(xlim = c(left, right), ylim = c(bottom, top)
)

ggsave(
  filename = "10x/figures/contour_plot_n1000.png", 
  contour_plot,
  width = 6480,
  height = 6480,
  dpi = 300,
  units = "px",
  device = "png",
  limitsize = FALSE
)

rm(contour_plot)
```