# Networked Contours

## Load General Libraries
```{r}
library(tidyverse)
library(sf)
library(ggmap)
library(tidygraph)
library(igraph)
```

## Read in Data
```{r}
# read in RDS
lines.sf <- readRDS("10x/variables/lines_sf.rds")
buses.sf <- readRDS("10x/variables/buses_sf.rds")
map <- readRDS("10x/variables/map_stadia_z14.rds")

# get the bounding box of map
left <- attr(map, "bb")$ll.lon
right <- attr(map, "bb")$ur.lon
bottom <- attr(map, "bb")$ll.lat
top <- attr(map, "bb")$ur.lat
```


# Conduct Network Analysis
- Calculate the distance matrix for the network --- yikes this might be too big
- Recalculate the voltages at each node
## Inspect the Input Data
```{r}
# reread in the data
lines.sf <- readRDS("10x/variables/lines_sf.rds")
buses.sf <- readRDS("10x/variables/buses_sf.rds")

# inspect the data
glimpse(buses.sf)
glimpse(lines.sf)

# look at values in lines.sf$name
# lines.sf$name %>% unique()
```

> glimpse(buses.sf)
Rows: 295,430
Columns: 5
$ name     <chr> "p13udm3166", "p13udm3168", "p13udm3169", "p13udm3170", "p13u…
$ voltage  <dbl> 0.9789937, 0.9798307, 0.9802534, 0.9706192, 0.9719708, 0.9730…
$ lng      <dbl> -122.2683, -122.2682, -122.2679, -122.2690, -122.2701, -122.2…
$ lat      <dbl> 37.86478, 37.86447, 37.86484, 37.86560, 37.86579, 37.86551, 3…
$ geometry <POINT [°]> POINT (-122.2683 37.86478), POINT (-122.2682 37.86447),…
> glimpse(lines.sf)
Rows: 295,672
Columns: 11
$ is_switch   <lgl> NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA…
$ is_open     <lgl> NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA…
$ is_fuse     <lgl> NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA…
$ kva         <dbl> 7, 10, 8, 10, 8, 8, 6, 10, 10, 8, 10, 8, 8, 6, 10, 8, 10, …
$ type        <fct> Transformer, Transformer, Transformer, Transformer, Transf…
$ length..km. <dbl> NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA…
$ ampacity    <dbl> NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA…
$ name        <chr> "tr(r:p13udt769-p13udt769lv)", "tr(r:p13udt770-p13udt770lv…
$ kvs         <list> <"12.47", "0.48">, <"7.2", "0.12">, <"7.2", "0.12">, <"7.…
$ phases      <list> <>, <>, <>, <>, <>, <>, <>, <>, <>, <>, <>, <>, <>, <>, <…
$ geometry    <POINT [°]> POINT (-122.2682 37.86477), POINT (-122.2691 37.8655…
> 

## Clean Up the Lines.SF and Buses.SF Data
```{r}
# create a unique numeric id for each bus
buses.sf <- buses.sf %>% mutate(id = as.numeric(factor(name)))

# create a unique numeric id for each line
lines.sf <- lines.sf %>% mutate(id = as.numeric(factor(name)))

# line names look like: "l(r:p13udm85893-p13udt22349lv)"
# the from bus is "p13udm85893" and the to bus is "p13udt22349lv"
# extract the from and to bus names from this string with regex
lines.sf$from_bus <- str_extract(lines.sf$name, "(?<=:)\\w+")
lines.sf$to_bus <- str_extract(lines.sf$name, "(?<=-)\\w+")
glimpse(lines.sf)

# now pull in the bus id for the from and to bus from buses.sf
lines.sf$from <- buses.sf$id[match(lines.sf$from_bus, buses.sf$name)]
lines.sf$to <- buses.sf$id[match(lines.sf$to_bus, buses.sf$name)]

# check for any empty lines$from or lines.sf$to
lines.sf.filt <- lines.sf %>%
  filter(!is.na(from) & !is.na(to))

# filter out any buses that are not to or from a line
buses.sf.filt <- buses.sf %>%
  filter(id %in% c(lines.sf.filt$from, lines.sf.filt$to))
```

## Build the Network
```{r}
# ensure all edges have a node
graph <- tidygraph::tbl_graph(
  nodes = buses.sf,
  edges = tibble(lines.sf.filt),
  directed = FALSE
)
saveRDS(graph, "10x/variables/graph.rds")
```

## Check the Number of Components in the Network
```{r}
# read in the network
graph <- readRDS("10x/variables/graph.rds")

# calculate components
graph <- graph %>%
  activate(nodes) %>%
  mutate(component = components(graph)$membership)

# add a column that counts the number of nodes in each component
graph <- graph %>%
  group_by(component) %>%
  mutate(n = n()) %>%
  ungroup()


graph
max(graph %>% activate(nodes) %>% as_tibble() %>% pull(component)) # 46k components, can't work with that

# count unique values of number of nodes in each component
graph %>% activate(nodes) %>% as_tibble() %>% count(component) %>% count(n)%>% print(n=50)

# show me a histogram of the number of nodes in each component
# use log scale
graph %>% activate(nodes) %>% as_tibble() %>% count(component) %>% ggplot(aes(x = n)) + geom_histogram(binwidth = 1) + scale_y_log10()


# add the "component" and "n" columns to buses.sf
buses.sf <- buses.sf %>%
  left_join(graph %>% activate(nodes) %>% as_tibble() %>% select(name, component, n), by = c("name" = "name"))

glimpse(buses.sf)

# keep just buses where n > 1
buses.sf.filt <- buses.sf %>%
  filter(n > 1000)

# plot the buses
buses.sf %>%
  ggplot() +
  geom_sf(aes(color = as.factor(component), size = 0.1)) +
  theme_minimal() +
  theme(legend.position = "none")

```
Without Capacitors: 46642 components
With Capacitors: 46820 components
Number of Capacitors: 178 > they don't join up the network, it's a parsing problem

> graph %>% activate(nodes) %>% as_tibble() %>% count(component) %>% count(n)%>% print(n=50)
Storing counts in `nn`, as `n` already present in input
ℹ Use `name = "new_name"` to pick a new name.
# A tibble: 23 × 2
        n    nn
    <int> <int>
 1      1 46600
 2      2    15
 3      3     2
 4      4     3
 5      5     1
 6      6     3
 7      7     1
 8      8     2
 9      9     1
10     18     1
11     35     1
12     44     1
13    220     1
14    456     1
15   1644     1
16   2042     1
17   3721     1
18   3931     1
19   6155     1
20   6316     1
21   6715     1
22  31875     1
23 185555     1

## Force Directed Network Layout FR
```{r}
# library(igraph) # for network analysis
# # read in graph
# graph <- readRDS("10x/variables/graph.rds")

# # create force directed layout
# fr <- as.data.frame(layout_with_fr(graph, dim = 2))
# ggplot(fr, aes(x = V1, y = V2)) +
#   geom_point(alpha = 0.2, size = 0.5) +
#   theme_bw()
```

## Run Network Analysis
```{r}
library(igraph) # for network analysis
# read in graph and inspect
graph <- readRDS("10x/variables/graph.rds")
graph
##############################################
# calculate the distances between all nodes
dist_matrix <- distances(graph, mode = "all")
##############################################
# calculate a similarity matrix as 1 / (h+1)^2
sim2_matrix <- 1 / ((dist_matrix + 1)^2)

##############################################
# export matrices
saveRDS(dist_matrix, "10x/variables/dist_matrix.rds")
saveRDS(sim2_matrix, "10x/variables/sim2_matrix.rds")

# remove variables from RAM until needed
rm(dist_matrix, sim2_matrix)
```

## Proceed with Weighted Network Analysis
```{r}
# import buses.sf
buses.sf <- readRDS("10x/variables/buses_sf.rds")
sim2_matrix <- readRDS("10x/variables/sim2_matrix.rds")

# set up a function to calculate weighted voltage considering n nearest neighbors by hops (including self)
neighbors <- 3

# function to weight voltage
weight_voltage <- function(voltage, row) {
  weights <- sim2_matrix[, row]
  indices <- order(weights, decreasing = TRUE)[1:neighbors]
  vw <- data.frame(
    voltages = buses.sf$voltage[indices],
    weights = weights[indices]
  )
  vw$weighted_voltage <- vw$voltages * vw$weights
  sum_weights <- sum(vw$weights)
  vw$scaled_weighted_voltage <- vw$weighted_voltage / sum_weights
  weighted_voltage <- sum(vw$scaled_weighted_voltage)
  return(weighted_voltage)
}
##############################################
# Apply weighted voltage function to buses.sf
Sys.time()
buses.sf <- buses.sf %>% mutate(
  weighted_voltage_n3_d2 = purrr::map_dbl(
    row_number(),
    ~ weight_voltage(voltage, .x)
  )
)
Sys.time()

# filter out any columns that contain "_dev_"
# buses.sf <- buses.sf %>% select(-contains("_dev_"))

# sort the columns by name
# buses.sf <- buses.sf %>% select(sort(names(.)))

# reexport buses.sf
saveRDS(buses.sf, "10x/variables/buses_sf.rds")
```



# Create Contours

## Load Contour Map Libraries
```{r}
library(terra)
library(gstat)
```

## Create Contours
```{r}
# load in lines buffer
buses.sf <- readRDS("10x/variables/buses_sf.rds")
lines.buffer <- readRDS("10x/variables/lines_buffer_100m.rds")

# Convert the lines buffer to a vector
lines.buffer.sv <- vect(lines.buffer)
# get the extent of lines.buffer.sv
ext <- ext(lines.buffer.sv)
# create a grid of the same extent
grd <- rast(ext, res=0.0001)

# convert buses.sf to buses.df
buses.df <- buses.sf %>% st_drop_geometry()
# drop nas from any column
buses.df <- na.omit(buses.df)
# rename lng and lat to x and y
names(buses.df) <- c("name", "voltage", "x", "y")
# glimpse(buses.df)

neighbors <- 1000
# define interpolation model
mg <- gstat(
  id = "idw", 
  formula = voltage~1, 
  locations = ~x+y, 
  data=buses.df, 
  nmax=neighbors, 
  set=list(idp = 2)
)
######################
contour <- interpolate(grd, mg, debug.level=0, index=1)
saveRDS(contour, "10x/variables/contour_n1000_rast.rds")
######################
# plot(contour)
contour.mask <- mask(contour, lines.buffer.sv)
saveRDS(contour.mask, "10x/variables/contour_mask_n1000_rast.rds")
# plot(contour.mask)

# export the contour to RDS
# class(contour) # [1] "SpatRaster", attr(,"package"), [1] "terra"

# clear out extra variables not needed later
rm(buses.sf, buses.df, lines.buffer, lines.buffer.sv, ext, grd, neighbors, mg)
```

## Prepare Contour for Plotting
```{r}
# read in contour object
contour.mask <- readRDS("10x/variables/contour_mask_n1000_rast.rds")

# convert to data.frame
contour.df <- as.data.frame(contour, xy=TRUE)

# rename idw.pred to voltage
contour.df <- contour.df %>% rename(voltage = idw.pred)

# remove rows with NA in any column
# contour.df <- na.omit(contour.df) # removes none

# export df 
saveRDS(contour.df, "10x/variables/contour_n1000_df.rds")
glimpse(contour.df)

rm(contour)
```
Rows: 3,553,912
Columns: 3
$ x        <dbl> -122.3447, -122.3446, -122.3445, -122.3444, -122.3443, -122.3…
$ y        <dbl> 37.90743, 37.90743, 37.90743, 37.90743, 37.90743, 37.90743, 3…
$ voltage <dbl> 0.9800596, 0.9800594, 0.9800592, 0.9800590, 0.9800588, 0.9800…

## Plot Contours
```{r}
# import map
map <- readRDS("10x/variables/map_stadia_z14.rds")
contour.df <- readRDS("10x/variables/contour_n1000_df.rds")

# glimpse(contour.df)
# plot it
contour_plot <- ggmap(map) +
  geom_raster(
    data = contour.df,
    aes(x=x, y=y, fill = voltage),
    alpha = 1.0,
    inherit.aes = FALSE
  ) +
#   geom_sf(
#     data = lines.sf, 
#     color="black", 
#     alpha = 0.05, 
#     inherit.aes = FALSE
#   ) +
  scale_fill_gradient2(
    low = "darkblue",
    mid = "white",
    high = "darkred",
    midpoint = 1,
    limits = c(0.95, 1.05),
    space = "Lab",
    guide = "colourbar",
    aesthetics = "fill",
    oob = scales::squish
  ) +
  theme(
    legend.position = c(0.96, 0.15),
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.x = element_blank(),
    axis.ticks.y = element_blank(),
    text = element_text(family = "Serif", size = 18)
  ) +
  coord_sf(xlim = c(left, right), ylim = c(bottom, top)
)

ggsave(
  filename = "10x/figures/contour_plot_n1000.png", 
  contour_plot,
  width = 6480,
  height = 6480,
  dpi = 300,
  units = "px",
  device = "png",
  limitsize = FALSE
)

rm(contour_plot)
```